---
title: "Download data"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:  
  html_document:
    toc: true
bibliography: refs.bib
---

<script>
$(document).ready(function(){
  var hide_divs = $("div.hideable");
  hide_divs.each(function(){
    // Wrap content in div
    $(this).wrapInner( "<div class='hideable_content', style='display: none;'></div>");
    
    // Add button
    $(this).prepend("<button id='toogle'>show</button>");
  });
  
  // Add hideable btn
  // Put the rest in a div
  
  $("div.hideable button#toogle").click(function(){
    var parent = $(this).parent();
    var target_div = $(parent).find("div.hideable_content");
    
    if(target_div.css("display") == "none"){
      target_div.show();
        $(this).text("Hide");
    } else {
      target_div.hide();
      $(this).text("Show");
    }
  });
});
</script>

```{r setup, include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, 
  fig.height = 4, fig.width = 7, dpi = 128, comment = "#R",
  cache.path = paste0(file.path("cache", "dl_from_wrds"), .Platform$file.sep), 
  fig.path   = paste0(file.path("fig"  , "dl_from_wrds"), .Platform$file.sep),
  error = FALSE)
options(digits = 4, scipen = 10, width = 70)
```


## Monthly data

The code below gets data from

- [`compm.fundq`](https://wrds-web.wharton.upenn.edu/wrds/tools/variable.cfm?library_id=130&file_id=66036).
- [`crsp.msf`](https://wrds-web.wharton.upenn.edu/wrds/tools/variable.cfm?library_id=137&file_id=67077).
- [`crsp.Ccmxpf_linktable`](https://wrds-web.wharton.upenn.edu/wrds/tools/variable.cfm?library_id=137&file_id=66976).
- [`crsp.msenames`](https://wrds-web.wharton.upenn.edu/wrds/tools/variable.cfm?library_id=137&file_id=67074).

I found [the following information](https://gist.github.com/iangow/583557b7b91a87ee1e545aa839ccbb8d) on the primary linker

> LINKPRIM clarifies the link's relationship to Compustat's marked primary security within the related range. "P" indicates a primary link marker, as identified by Compustat in monthly security data. "C" indicates a primary link marker, as identified by CRSP to resolve ranges of overlapping or missing primary markers from Compustat in order to produce one primary security throughout the company history. "J" indicates a joiner secondary issue of a company, identified by Compustat in monthly security data.

I also found [this](http://www.kaikaichen.com/?p=138) which is also shown 
[here](http://www.wrds.us/index.php/forum_wrds/viewthread/609/)

> As of the February 2014 release, USEDFLAG is no longer used in the WRDS CCM web queries.  Please select LINKTYPES LC, LU, and LS for the same results. These represent the vast majority of the links between CRSP securities and Compustat companies, without introducing duplicate data.
>
> The WRDS-created linking dataset (ccmxpf_linktable) has been deprecated. It will continue to be created for a transition period of 1 year. SAS programmers should use the Link History dataset (ccmxpf_lnkhist) from CRSP.

Further from the "CRSP/COMPUSTAT merged database guide"

> LNKTYPE: Link type code. Each link is given a code describing the connection 
> between the CRSP and Compustat data. Values are:
> 
> LC – Link research complete. Standard connection between databases
>
> LU – Unresearched link to issue by CUSIP
>
> LS – Link valid for this security only. Other CRSP PERMNOs with the same 
> PERMCO will link to other GVKEYs
> 
> LINKPRIM: Primary issue marker for the link. Based on
> Compustat Primary/Joiner flag (PRIMISS), indicating
> whether this link is to Compustat’s marked primary
> security during this range. 
>
> P = Primary, identified by Compustat in monthly security data.
> 
> J = Joiner secondary issue of a company, identified by Compustat in monthly 
> security data.
>
> C = Primary, assigned by CRSP to resolve ranges of overlapping or missing 
> primary markers from Compustat in order to produce one primary security
> throughout the company history.

Ideas

 - do as in [merge_funda_crsp_bycusip.sas](https://wrds-www.wharton.upenn.edu/pages/support/research-wrds/sample-programs/program-merge-crsp-and-compustat-using-cusip/) but with compm.fundq
 - figure out whether this guide is legit and use it 
   http://www.ruidaiwrds.info/data/linking-crsp-and-compustat

See also [crspmerge.sas](https://wrds-web.wharton.upenn.edu/wrds/support/code_show.cfm?path=CRSP/crspmerge.sas)
on how to merge `crsp.msf` and `crsp.msenames`. 

<div class="hideable">

```{r get_monthly_data}
overwrite <- FALSE
wrds <- .wrdsconnect()
for(yr in 1962:2017){
  sql. <- paste(readLines("sql/merged_monthly.sql"), collapse = "\n")
  f_name <- file.path("data", "CRSP", "monthly_stock", paste0(yr, ".RDS"))
  
  if(file.exists(f_name) && !overwrite){
    old_sql <- readRDS(f_name)$sql
    if(isTRUE(all.equal(old_sql, sql.))){
      cat(sQuote(f_name), "exists and will not be overwritten...\n")
      next
    } else
      cat("Old", sQuote("cmds"), "object does not match new. Running new\n")
  }
  
  params <- list(paste0(yr, "-12-31"), paste0(yr, "-01-01"))
  cat("Getting data...\n")
  sql. <- paste(readLines("sql/merged_monthly.sql"), collapse = "\n")
  cat("\nRunning\n\n\t", gsub("\n", "\n\t", sql.), 
      "\n\nwith parameters: ", 
      sQuote(paste(deparse(params), sep = "\n")), sep = "")
  
  res <- dbSendQuery(wrds, sql.)
  res <- dbBind(res, params)
  dat <- dbFetch(res, n = -1)
  dbClearResult(res)
  
  # checks 
  stopifnot(max(xtabs(~ cusip, dat)) == 12L, any(xtabs(~ cusip, dat) == 12L))
  stopifnot(max(xtabs(~ gvkey, dat)) == 12L, any(xtabs(~ gvkey, dat) == 12L))
  
  # save
  cat("Final number of rows are ", nrow(dat), ". Saving data...\n", sep = "")
  saveRDS(list(data = dat, sql = sql.), f_name)
}
```

</div>

### Stack monthly data

```{r stack_month}
# where we want data after
yr_lb <- 1978L

#####
# load data
file_stack <- "data/CRSP/monthly_stacked.RDS"

fs <- list.files("data/CRSP/monthly_stock/", full.names = TRUE)
fs <- fs[grepl("RDS$", fs)]
length(fs) # number of files

# only take data after a given year
yr <- as.integer(gsub("(.+/)(\\d{4})([^/]+)$", "\\2", fs))
fs <- fs[yr > yr_lb]
length(fs) # number of files

# find last edit and check if we need to update 
max_ti <- local({
  ti <- lapply(lapply(fs, file.info), "[[", "mtime")
  ti[[which.max(unlist(ti))]]
})

do_stack <- 
  !file.exists(file_stack) || file.info(file_stack)$mtime <= max_ti

if(do_stack){
  cat("Stacking data\n")
  
  # stack data and save
  stack_dat <- do.call(rbind, lapply(fs, function(x) readRDS(x)$data))
  saveRDS(stack_dat, file_stack)
  
} else 
  cat("Not stacking data and using previous file\n")
```

```{r cleanup_month_crsp, echo = FALSE}
rm(list = ls())
wrds <- .wrdsconnect()
```

### Multiple permnos in the same month 

As mentioned on the 
[Fama-French Factors site on WRDS](https://wrds-www.wharton.upenn.edu/pages/support/applications/risk-factors-and-industry-benchmarks/fama-french-factors/), 
some permcos will match with multiple permnos. We can

1. handle this in the same way as in the above link. I.e. add the market value up on permco
   level and then add the computed market value  to the permno with the highest market value.
2. disregard this assuming there are few firms with multiple permnos. 

We make the following query to check the latter

<div class="hideable">

```{r how_may_mult_permnos, cache = 1}
#####
# find number of permco with x number of permnos for each month
sql. <- "
select b.ym, b.n_permno, count(distinct b.permco) as n_firms
from (
  select a.ym, a.permco, count(distinct a.permno) as n_permno
  from (
    select date_trunc('month', date) as ym, permco, permno
    from crsp.msf) as a 
  group by a.ym, a.permco) as b
group by b.ym, b.n_permno"

que <- dbSendQuery(wrds, sql.)
n_permnos <- dbFetch(que, n = -1)
dbClearResult(que)

n_permnos$n_permno <- as.integer(n_permnos$n_permno)
n_permnos$n_firms <- as.integer(n_permnos$n_firms)

library(data.table)
n_permnos <- data.table(n_permnos)

# some have quite a few
max(n_permnos$n_permno)

#####
# plot
setkey(n_permnos, ym, n_permno)
tmp <- n_permnos[, .(
  `1`      =     n_firms[n_permno == 1L]     / sum(n_firms), 
  `2`      =     n_firms[n_permno == 2L ]    / sum(n_firms), 
  `[3, 5]` = sum(n_firms[n_permno %in% 3:5]) / sum(n_firms), 
  `>5`     = sum(n_firms[n_permno > 5]     ) / sum(n_firms)), 
  by = ym]

par(mar = c(5, 4, .5, .5))
matplot(tmp$ym, tmp[, -1L, with = FALSE], type = "l", lty = 1, 
        ylab = "Fraction of permco by # permno", xlab = "Time")
abline(h = 1, lty = 2)
legend("topleft", legend = colnames(tmp)[-1], lty = 1, col = 1:4, bty = "n")

matplot(tmp$ym, tmp[, -(1:2), with = FALSE], type = "l", lty = 1, 
        ylab = "Fraction of permco by # permno", xlab = "Time")
abline(h = 0, lty = 2)
legend("topleft", legend = colnames(tmp)[-(1:2)], lty = 1, col = 1:3, bty = "n")
```

</div>

We will later do as in 1. above.

### Delisting data

Download delisting dates, their codes, and add the first month in which we see 
a return

<div class="hideable">

```{r delist_data}
f <- file.path("data", "CRSP", "delist.RDS")

sql. <- "
   select a.permno, a.permco, a.dlstdt, a.dlstcd, a.nwperm, a.nwcomp, b.first_ret
   from crsp.msedelist as a
   left join (
    select min(date) as first_ret, permno
    from crsp.msf
    group by permno
   ) as b on a.permno = b.permno
   /* code 1xx is active status */
   where a.dlstdt > '1979-01-01' and a.dlstcd > 199" 

if(!file.exists(f) || !isTRUE((dat <- readRDS(f))$sql == sql.)){
  cat(sQuote(f), "not found\n")
  
  res <- dbSendQuery(wrds, sql.)
  dat <- dbFetch(res, n = -1)
  dbClearResult(res)
  
  # sanity check
  stopifnot(
    # some only have one date and this is where they delist?
    any(dat$dlstdt == dat$first_ret),
    all(dat$dlstdt >= dat$first_ret), 
    !anyDuplicated(dat$permno),
    anyDuplicated(dat$permco) > 0)
  
  dat <- list(data = dat, sql = sql.)
  saveRDS(dat, f)
  
} else
  cat(sQuote(f), "found\n")

table(dat$data$dlstcd)
length(unique(dat$data$permno))
length(unique(dat$data$permco))
```

</div> 

`DLSTDT` is the delisting day and `DLSTCD` is the delisting code. It can these 
first digit values

| Code | Category                      |
|------|-------------------------------|
| 100  | Active                        |
| 200  | Mergers                       |
| 300  | Exchanges                     |
| 400  | Liquidations                  |
| 500  | Dropped                       |
| 600  | Expirations                   |
| 900  | Domestics that became Foreign |

Further

> The second and third digits of the delisting codes provide further detail of 
delisting events.

It seems important to note that

> Delisting Date
> 
> Delisting Date is an integer containing the date in YYYYMMDD format of a 
security's last price on the current exchange. If the security is still active, 
Delisting Date is set to the last date of available price data. Delisting date 
is never missing.

## Daily data

The code below gets data from

- [`crsp.dsf`](https://wrds-web.wharton.upenn.edu/wrds/tools/variable.cfm?library_id=137&file_id=67061).
- [`crsp.dse`](https://wrds-web.wharton.upenn.edu/wrds/tools/variable.cfm?library_id=137&file_id=67053).
- [`crsp.dsenames`](https://wrds-web.wharton.upenn.edu/wrds/tools/variable.cfm?library_id=137&file_id=67058).

<div class="hideable">

```{r dl_crsp}
overwrite <- FALSE
vals <- expand.grid(year = 1978:2017, quater = 1:4)

for(i in 1:nrow(vals)){
  #####
  # setup SQL queries
  y <- vals[i, "year"]
  q <- vals[i, "quater"]
  
  cat("Getting data...\n")
  sql. <- paste(readLines("sql\\daily_crsp.sql"), collapse = "\n")
  
  start_date <- paste0(y, "-", sprintf("%02d", 1L + (q - 1L) * 3L), "-01")
  sql. <- gsub("\\{STAT_DATE\\}", start_date, sql.)
  
  end_date <- paste0(
    y, "-", sprintf("%02d", q * 3L), 
    "-", switch (q, "1" = 31L, "2" = 30L, "3" = 30L, "4" = 31L))
  sql. <- gsub("\\{END_DATE\\}", end_date, sql.)
  
  sql. <- gsub("\\{S\\}", "d", sql.)
  cmds <- strsplit(sql., ";")[[1]]
  
  f_name <- file.path("data", "CRSP", "daily_stock", paste0(y, "_Q", q, ".RDS"))
  if(file.exists(f_name) && !overwrite){
    old_cmds <- readRDS(f_name)$cmds
    if(isTRUE(all.equal(old_cmds, cmds))){
      cat(sQuote(f_name), "exists and will not be overwritten...\n")
      next
    } else
      cat("Old", sQuote("cmds"), "object does not match new. Running new\n")
    
  }
  
  out <- lapply(cmds, function(cmd){
      require(RPostgres)
      cmd <- paste0(cmd, ";")
      # cat("Running the following sql script\n")
      # cat(cmd, "\n")
      
      res <- dbSendQuery(wrds, cmd)
      out <- dbFetch(res, n = -1)
      cat("The object requires", object.size(out) / 10^6, " mb\n")
      
      # clean-up
      dbClearResult(res)
      dbDisconnect(wrds)
      .GlobalEnv$wrds <- .wrdsconnect()
      
      out
    })
  
  cat("Number of rows for the result is\n")
  print(lapply(out, nrow))
  
  cat("Merging...\n")
  stock_dat <- out[[1]]
  event_dat <- out[[2]]
  rm(out)
  
  stock_dat$is_stock <- TRUE
  event_dat$is_event <- TRUE
  
  out <- merge(stock_dat, event_dat, by = c("permno", "date"), all = TRUE)
  rm(stock_dat, event_dat)
  
  cat("Carry last value forward...\n")
  stopifnot(!is.unsorted(order(out$permno, out$date)))
  require(zoo)
  out$exchcd <- 
    do.call(c, tapply(out$exchcd, out$permno, na.locf, na.rm = FALSE))
  
  cat("Removing non-stock data rows. There are", nrow(out), "rows before...\n")
  out <- out[!is.na(out$is_stock), ]
  out[c("is_stock", "is_event")] <- NULL
  
  cat("There are", nrow(out), "rows after\n")
  
  cat("Droppings rows which we now are not listed on the relevant exchanges...\n")
  print(xtabs(~out$exchcd, addNA = TRUE))
  
  out$keep <- is.na(out$exchcd) | out$exchcd %in% c(-(2:1), 1:3)
  out <- out[out$keep, ]
  out$keep <- NULL
  
  cat("Final number of rows are ", nrow(out), ". Saving data...\n", sep = "")
  saveRDS(list(data = out, cmds = cmds), f_name)
}
```

</div>

```{r look_at_db_crsp, echo = FALSE, eval = FALSE}
res <- dbSendQuery(wrds, "SELECT DISTINCT HEXCD FROM CRSP.DSF;")
dbFetch(res, n = -1)
#   HEXCD
# 1     1
# 2     2
# 3     3
# 4     4
dbClearResult(res)

res <- dbSendQuery(wrds, "SELECT DISTINCT EXCHCD FROM CRSP.DSE;")
sort(dbFetch(res, n = -1)[[1]])
# [1] -2 -1  0  1  2  3  4 10 13 16 17 19 20 31 32 33
dbClearResult(res)

# all data sets
res <- dbSendQuery(wrds, "select distinct table_name
                   from information_schema.columns
                   where table_schema='comp'
                   order by table_name")
dbFetch(res, n=-1)
dbClearResult(res)

# columns in data set
res <- dbSendQuery(wrds, "select column_name
                   from information_schema.columns
                   where table_schema='comp'
                   and table_name='funda'
                   order by column_name")
dbFetch(res, n=-1)
dbClearResult(res)
```

### Stack daily data

Create list of `PERMNO`s to include from the merged COMPUSTAT and CRSP data

```{r unique_permno}
uni_perm <- unique(readRDS("data/CRSP/monthly_stacked.RDS")$permno)
```

First we merge the data. We start by stacking the daily data

```{r merge_dat_daily}
# where we want data after
yr_lb <- 1978L

#####
# load data
file_stack <- "data/CRSP/daily_stacked.RDS"

fs <- list.files("data/CRSP/daily_stock/", full.names = TRUE)
fs <- fs[grepl("RDS$", fs)]
length(fs) # number of files

# only take data after a given year
yr <- as.integer(gsub("(.+/)(\\d{4})([^/]+)$", "\\2", fs))
fs <- fs[yr > yr_lb]
length(fs) # number of files

# find last edit and check if we need to update 
max_ti <- local({
  ti <- lapply(lapply(fs, file.info), "[[", "mtime")
  ti[[which.max(unlist(ti))]]
})

do_stack <- 
  !file.exists(file_stack) || file.info(file_stack)$mtime <= max_ti

if(do_stack){
  cat("Stacking data\n")
  
  library(data.table)
  # stack data and save
  stack_dat <- do.call(rbind, lapply(fs, function(x){
    new_dat <- readRDS(x)$data
    
    # compute market value and add sum of market value to the permno within a 
    # (permco, date) pair with the highest mv
    new_dat <- data.table(new_dat)
    setkey(new_dat, permco, date)
    new_dat[, mv := abs(prc) * shrout][
      , `:=`(mv_use = mv, has_more = .N > 1L & sum(!is.na(mv)) > 1L), 
      by = .(permco, date)][
        has_more == TRUE, mv_use := 
          ifelse(1:.N == which.max(mv), sum(mv, na.rm = TRUE), mv), 
        by = .(permco, date)]
      
    cat(with(new_dat, mean(!is.na(mv) & mv != mv_use)), " has a new market", 
        "value\n")
    
    stopifnot(all(new_dat[, is.na(mv) == is.na(mv_use) & 
                            (is.na(mv) | mv <= mv_use)]))
    new_dat[, mv := mv_use]
    new_dat <- as.data.frame(new_dat)
    
    # only keep the rows we need
    new_dat <- subset(new_dat, permno %in% uni_perm)
    
    # only keep the columns we need and adjust some types
    new_dat$permno <- as.integer(new_dat$permno + sign(new_dat$permno) * 1e-8)
    new_dat[, c("permno", "date", "mv", "ret")]
    
  }))
  saveRDS(stack_dat, file_stack)
  
} else 
  cat("Not stacking data and using previous file\n")
```


## Treasury Bill

```{r cleanup_bill, echo = FALSE}
rm(list = ls())
wrds <- .wrdsconnect()
```

We need the 1-year Treasury Bill to compute the distance-to-default as in e.g., 
@Vassalou04. We can get this from St. Louis Federal Reserve Bank's. See 
[this site](https://fred.stlouisfed.org/series/GS1). We can also get 
this from [frb.rates_daily](https://wrds-web.wharton.upenn.edu/wrds/tools/variable.cfm?library_id=34&file_id=53609).

<div class="hideable">

```{r get_tres, cache = TRUE, message = FALSE}
#####
# dara from St. Louis Federal Reserve Bank's
library(quantmod)
library(RPostgres)
getSymbols('DGS1', src = 'FRED')
DGS1 <- data.frame(date = index(DGS1), r1y = drop(coredata(DGS1)) / 100)
getSymbols('DGS3MO', src = 'FRED')
DGS3MO <- data.frame(date = index(DGS3MO), r3m = drop(coredata(DGS3MO)) / 100)
fred_dat <- merge(DGS1, DGS3MO, all = TRUE)

#####
# data from WRDS 
sql. <- "
select date, 
  TCMNOM_M3 / 100 as m3 /* Treasury Constant Maturity 3-month */, 
  TCMNOM_Y1 / 100 as y1 /* Treasury Constant Maturity 1-year */ 
  from frb.RATES_DAILY"

res <- dbSendQuery(wrds, sql.)
dat <- dbFetch(res, n = -1)
dbClearResult(res)

#####
# compare the two
tmp <- merge(fred_dat, dat, by = "date", all = TRUE)
matplot(tmp$date, tmp[, c("r1y", "y1")], pch = 16)
matplot(tmp$date, tmp[, c("r3m", "m3")], pch = 16)

# save the series
saveRDS(fred_dat, file.path("data", "tress_bill.RDS"))
```

</div>

## Inflation index
Download inflation index from the 
[mcti](https://wrds-web.wharton.upenn.edu/wrds/tools/variable.cfm?library_id=138&file_id=67186)
database. See 
[crsp.com/products/documentation/crsp-treasury-and-inflation-indexes-0](http://www.crsp.com/products/documentation/crsp-treasury-and-inflation-indexes-0). 

```{r get_inflation}
res <- dbSendQuery(
  wrds, "select caldt, cpiind from crsp.mcti")
dat <- dbFetch(res, n = -1)
dbClearResult(res)

# December 29, 1972 (19721229) equals 100
dat$caldt <- as.Date.character(dat$caldt, "%Y%m%d")
with(dat, plot(caldt, cpiind, type = "l"))
abline(h = 100, lty = 2) 
abline(v = as.Date("1972-12-29"), lty = 2)

saveRDS(dat, file.path("data", "inflation.RDS"))
```

## Stock indices

Get S&P500 index from [crspa.dsp500](https://wrds-web.wharton.upenn.edu/wrds/tools/variable.cfm?library_id=137&file_id=67002)

<div class="hideable">

```{r SP500}
sql. <- "
select caldt, 
  ewretd, /* Equal-Weighted Return (includes distributions) */ 
  ewretx, /* Equal-Weighted Return (excluding dividends) */ 
  vwretd, /* Value-Weighted Return (includes distributions) */ 
  vwretx, /* Value-Weighted Return (excluding dividends) */ 
  totval /* Total Market Value */ 
  from crspa.dsp500"

res <- dbSendQuery(wrds, sql.)
sp_dat <- dbFetch(res, n = -1)
dbClearResult(res)

sp_dat <- subset(sp_dat, caldt >= as.Date("1960-01-01"))
plot(sp_dat$vwretd ~ sp_dat$caldt, type = "h") # value weighted
plot(sp_dat$ewretd ~ sp_dat$caldt, type = "h") # equal weighted

saveRDS(sp_dat, file.path("data", "SP500.RDS"))
```

</div>

Get AMEX and NYSE indices. I think we can use the data from [crspa.dsic](https://wrds-web.wharton.upenn.edu/wrds/tools/variable.cfm?library_id=137&file_id=66997) (the title is *(Daily) NYSE/AMEX Capitalization Deciles, Annual Rebalanced*). 
I think it is the *CRSP Market Indexes* mentioned [here](http://www.crsp.com/products/documentation/stock-file-indexes-0). 

<div class="hideable">

```{r marked_indices}
sql. <- "
select caldt, 
  vwretd, /* Value-Weighted Return-incl. dividends */ 
  vwindd, /* Index Level Associated with VWRETD */ 
  vwretx, /* Value-Weighted Return-excl. dividends */ 
  vwindx, /* Index Level Associated with VWRETX */ 
  ewretd, /* Equal-Weighted Return-incl. dividends */
  ewindd, /* Index Level Associated with EWRETD */
  ewretx, /* Equal-Weighted Return-excl. dividends */
  ewindx, /* Index Level Associated with EWRETX */
  sprtrn, /* Return on the S&P 500 Index */  
  spindx, /* Level of the S&P 500 Index */
  totval, /* Total Market Value */ 
  totcnt  /* Total Market Count */
  from crspa.dsic"

res <- dbSendQuery(wrds, sql.)
idx_dat <- dbFetch(res, n = -1)
dbClearResult(res)

head(idx_dat)
plot(idx_dat$vwretd ~ idx_dat$caldt, type = "h")
plot(idx_dat$sprtrn ~ idx_dat$caldt, type = "h")
plot( sp_dat$vwretd ~ sp_dat$caldt , type = "h", xlim = range(idx_dat$caldt))

saveRDS(idx_dat, file.path("data", "index.RDS"))
```

</div>

We can also make the index our-self 

<div class="hideable">

```{r own_idx}
f_name <- file.path("data", "index_self.RDS")
.sql <- paste(readLines("sql/index.sql"), collapse = "\n")

do_rerun <- !file.exists(f_name) || readRDS(f_name)$sql != .sql
if(do_rerun){
  cat("Getting data\n")
  res <- dbSendQuery(wrds, .sql)
  idx_self_dat <- dbFetch(res, n = -1)
  dbClearResult(res)
  
  # save result
  saveRDS(list(data = idx_self_dat, sql = .sql), f_name)
  
} else{
  cat("Loading data\n")
  idx_self_dat <- readRDS(f_name)$data
  
}

#####
# check data
nrow(idx_dat)
nrow(idx_self_dat)
tmp <- merge(idx_dat, idx_self_dat, by.x = "caldt", by.y = "date", 
             suffixes = c("_sic", "_self"))

# results should be the same pre NASDAQ
with(subset(tmp, caldt <= as.Date("1970-01-01")), 
     cor(vwretd    , vwret))
with(subset(tmp, caldt <= as.Date("1970-01-01")), 
     cor(ewretx_sic, ewretx_self))
with(subset(tmp, caldt <= as.Date("1970-01-01")), 
     cor(totval    , market_size))
with(subset(tmp, caldt <= as.Date("1970-01-01")), 
     cor(totcnt    , n_firms))

with(subset(tmp, caldt <= as.Date("1970-01-01")),
     plot(caldt, (totval -  market_size) / totval, 
          type = "l", main = "pre"))
abline(h = 0, lty = 2)

# may differ due to NASDAQ post NASDAQ starts
with(tmp, cor(vwretd    , vwret))
with(tmp, cor(ewretx_sic, ewretx_self))
with(tmp, cor(totval    , market_size))
with(tmp, cor(totcnt    , n_firms))

with(tmp, plot(caldt, (totval -  market_size) / market_size, 
               type = "l", main = "post"))
abline(h = 0, lty = 2)
with(tmp, plot(caldt, n_firms, type = "l", main = "post"))
with(tmp, plot(caldt, vwretd - vwret, type = "h", main = "post"))
abline(h = 0, lty = 2)
with(tmp, plot(caldt, market_size, type = "l", main = "market size", 
               log = "y"))
```

</div>

## Compustat

We start by looking at what variables people have used before in order to gather
what we need to download. @Shumway01 uses (some of those below are parts of 
ratios; the codes in table 1 refer to the data item number is the [annual data](http://www.crsp.com/products/documentation/annual-data-industrial))

- `WCAPQ`: Working Capital (Balance Sheet)
- `ATQ`: Assets - Total
- `REQ`: Retained Earnings
- `OIADPQ`: Operating Income After Depreciation (EBIT)
- `SALEQ`: Sales/Turnover (Net)
- `NIQ`: Net Income (Loss)
- `LTQ`: Liabilities - Total
- `LCTQ`: Current Liabilities - Total
- `ACTQ`: Current Assets - Total
- The logarithm of each firm’s size relative to the total size of the NYSE and 
  AMEX market is used
- Firm’s past excess return in year $t$ as the return of the firm in year 
  $t - 1$ minus the value-weighted CRSP NYSE/AMEX index return in year $t - 1$
- The idiosyncratic standard deviation of each firm’s stock returns
   
@Chava04 use 

- Net income to total assets 
- Total liabilities to total assets
- Relative size defined as the logarithm of each firm’s equity value 
  divided by the total NYSE/AMEX market equity value
- A dummy for ten industries by the 4-digit SIC code with some interactions. 
  The four digit SIC codes for each company is from CRSP and it is 
  time-varying. 
- Excess return defined as the monthly return of the firm minus the 
  value-weighted CRSP NYSE/AMEX index return cumulated to
  obtain the yearly return
- The stock’s volatility. The stock’s volatility for the present year is 
  computed as the sample standard deviation using the last sixty observable 
  daily market prices
  
The latter is not what @Shumway01 does.

@Duffie07 use (see [this site](http://www.crsp.com/products/documentation/quarterly-data-industrial) for the legacy codes they refer to)

- `DLCQ`: Debt in Current Liabilities (data45)
- `LCTQ`: Current Liabilities - Total (data49)
- `DLTTQ`: Long-Term Debt - Total (data51)
- `CSHOQ`: Common Shares Outstanding (data61)
- Distance-to-default

@Campbell08 use (I think I have the right variables)

- Net Income to Market-valued Total Assets
- Net Income to Total Assets
- Total liabilities relative to total assets
- Total liabilities divided by the sum of market equity and book liabilities
- Cash and short-term assets, `CHEQ`, to the market value
- Market-to-book ratio (Book Value = Total Assets – Total Liabilities). 
  Seems like we can compute with `sum(SEQQ, coalesce(TXDITC,sum(TXDB, ITCB)),-pstk_new)`
  were `SEQQ` = Shareholders' Equity - Total, 
  `TXDITCQ` = Deferred Taxes and Investment Tax Credit, `TXDB` and `ITCB` are 
  only the yearly data, and 
  `pstk_new=coalesce(PSTKRV,PSTKL,PSTK); /*preferred stock*/`
- Log excess return on each firm’s equity relative to the S&P 500 index
- Relative size of each firm measured as the log ratio of its market 
  capitalization to that of the S&P 500 index
- Firm’s log price per share, truncated above at $15

@Bharath08 use

- Various forms of distance-to-default
- `DLCQ`: Debt in Current Liabilities (data45)
- `DLTTQ`: Long-Term Debt - Total (data51)
- Net income to total assets
- Each firm's past excess return in year $t$ as the return of the firm in year 
  $t - 1$ minus the value- weighted CRSP NYSE/AMEX index return in return in 
  year $t - 1$
- Each firm's annual returns are calculated by cumulating monthly returns. 

@Lando10 use 

- 1-year equity return
- 1-year distance-to-default
- Quick ratio, calculated as the sum of cash, short-term investments and total 
  receivables divided by current liabilities
- Percentage short term debt, calculated as debt in current liabilities 
  divided by the sum of debt in current liabilities and long-term debt
- Book asset value (log)

@Chava11 use

- Four digit `DNUM` industry codes from COMPUSTAT
- A industry dummy which is one if the median stock return in that industry during
the year is less than -20%
- The relative size, defined as the logarithm of each firm's equity value 
divided by the total NYSE/AMEX/NASDAQ market capitalization
- The excess return, defined as the return of the firm minus the 
value-weighted CRSP NYSE/AMEX/NASDAQ index return
- The ratio of net income to total assets
- The ratio  of total liabilities to  total assets
- The volatility, defined as the idiosyncratic standard deviation of the 
firm's monthly stock returns
- The firm's trailing one year stock return, computed by cumulating the firm's
  monthly return
- The distance-to-default constructed as in @Bharath08
- The logarithm of the total assets
- The market-to-book ratio
- The ratio of property plant and equipment to total assets

@Duan12 use

- Distance-to-default
- Sum of cash and short-term investments to the total assets
- Net income to the total assets.
- Logarithm of the ratio of a firm’s market equity value to the average market 
  equity value of the S&P500 firms
- Market-to-book asset ratio
- 1-year idiosyncratic volatility as in @Shumway01

@Lando13 use (I think I have the right variables)

- Quick ratio: book value of cash and short-term investments, `CHEQ`, added the book 
  value of total receivables, `RECTQ`, divided by  the book value of total current 
  liabilities, `LCTQ`
- Pledgeable assets: book value of total current assets, `LCTQ`, plus the book value of 
  net property, plant, and equipment 
- Trailing 1-year equity return
- Trailing 1-year distance-to-default
- Percentage short-term debt:  the book value of debt in current liabilities 
  divided by the sum of the book value of debt in current liabilities and the 
  book value of total long-term debt
  
Maybe check [WRDS Financial Ratios Suite](https://wrds-www.wharton.upenn.edu/pages/support/research-wrds/sample-programs/wrds-financial-ratios-suite/).
We get the data from [`compm.fundq`](https://wrds-web.wharton.upenn.edu/wrds/tools/variable.cfm?library_id=130&file_id=66036). The quarterly does not have a SIC code

```{r issue_w_sic}
#####
# quaterly data
res <- dbSendQuery(wrds, "select column_name
                   from information_schema.columns
                   where table_schema='comp'
                   and table_name='fundq'
                   order by column_name")
out <- dbFetch(res, n=-1)$column_name
out[grepl("^si", out)]
dbClearResult(res)

#####
# yearly data
res <- dbSendQuery(wrds, "select column_name
                   from information_schema.columns
                   where table_schema='comp'
                   and table_name='funda'
                   order by column_name")
out <- dbFetch(res, n=-1)$column_name
out[grepl("^si", out)]
dbClearResult(res)
```

So we need the [`compm.funda`](https://wrds-web.wharton.upenn.edu/wrds/tools/variable.cfm?library_id=130&file_id=66034)
which we anyway do to also get the annual figures.

```{r get_compustat}
file_comp <- "data/compustat.RDS"

#####
# we form the sql and check if we need update the data
sql. <- paste(readLines("sql/compustat.sql"), collapse = "\n")

# get parameters for query
gvkeys_keep <- unique(readRDS("data/CRSP/monthly_stacked.RDS")$gvkey)

res <- dbSendQuery(wrds, "select distinct gvkey from compm.fundq")
gvkeys_db <- dbFetch(res, n = -1)$gvkey
dbClearResult(res)  
excl_gvkey <- setdiff(gvkeys_db, gvkeys_keep)

sql. <- gsub(
  "\\{GVKEY_EXL\\}", paste0("(", paste0("'", excl_gvkey, "'", collapse = ", "), 
                            ")"), 
  sql.)
sql. <- gsub("\\{MIN_DATE\\}", "1978-01-01", sql.)

# check if we need to update and if we do then do it
do_update <- !file.exists(file_comp) || readRDS(file_comp)$sql != sql.

if(do_update){
  cat("Updating file\n")
  
  # send query, get data, and save
  res <- dbSendQuery(wrds, sql.)
  dat <- dbFetch(res, n = -1)
  dbClearResult(res)
  
  saveRDS(list(data = dat, sql = sql.), file_comp)
} else
  cat("Not updating file\n")
```


## References
