---
title: "Market data"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:  
  html_document:
    toc: true
bibliography: refs.bib
---

<script>
$(document).ready(function(){
  var hide_divs = $("div.hideable");
  hide_divs.each(function(){
    // Wrap content in div
    $(this).wrapInner( "<div class='hideable_content', style='display: none;'></div>");
    
    // Add button
    $(this).prepend("<button id='toogle'>show</button>");
  });
  
  // Add hideable btn
  // Put the rest in a div
  
  $("div.hideable button#toogle").click(function(){
    var parent = $(this).parent();
    var target_div = $(parent).find("div.hideable_content");
    
    if(target_div.css("display") == "none"){
      target_div.show();
        $(this).text("Hide");
    } else {
      target_div.hide();
      $(this).text("Show");
    }
  });
});
</script>

## Load data

```{r static_setup, include=FALSE, cache=FALSE}
# please do not set options here that could change...
knitr::opts_chunk$set(
  cache.path = 
    paste0(file.path("cache", "market-data"), .Platform$file.sep), 
  fig.path = 
    paste0(file.path("fig"  , "market-data"), .Platform$file.sep))
```

```{r def_data_files}
# assign file names
fs <- list(
  daily    = file.path("data", "CRSP", "daily_stacked.RDS"), 
  monthly  = file.path("data", "CRSP", "monthly_stacked.RDS"),
  index    = file.path("data", "index_self.RDS"),
  index_o  = file.path("data", "index.RDS"),
  ym_R     = file.path("R", "ym.R"), 
  map_file = file.path("data", "moody-compu-link.RDS"))
```

```{r check_rebuild, echo = FALSE, cache = TRUE, cache.extra = tools::md5sum(unlist(fs))}
# see https://stackoverflow.com/a/52163751/5861244
knitr::opts_chunk$set(cache.rebuild = TRUE)
```

```{r setup, include=FALSE, cache=FALSE}
# please do set options here that could change...
knitr::opts_chunk$set(
  echo = TRUE, fig.height = 4, fig.width = 7, dpi = 128, comment = "#R", 
  error = FALSE)
options(digits = 4, scipen = 10, width = 70)
```

```{r load_dat}
daily <- readRDS(fs$daily)
monthly <- readRDS(fs$monthly)
```

Keep only the permno we need 

```{r keep_needed}
# Get the list of permnos and permco to keep
map_obj <- readRDS(fs$map_file)
gvkey_keep <- unique(map_obj$gvkey)
permno_keep <- with(monthly, unique(permno[gvkey %in% gvkey_keep]))

# subset 
monthly <- subset(monthly, permno %in% permno_keep)
daily   <- subset(daily  , permno %in% permno_keep)

rm(map_obj, monthly)
```

Make year month variables

```{r source_ym}
source(fs$ym_R)
daily$ym <- make_ym(daily$date)
```

Load the index data

```{r load_idx}
idx <- readRDS(fs$index)$data

idx_o <- readRDS(fs$index_o)
colnames(idx_o)[colnames(idx_o) %in% c("caldt", "vwretd")] <- 
  c("date", "vwret")
stopifnot(all(idx$date %in% idx_o$date))

# keep only the columns we need
idx <- merge(idx[, c("date", "market_size")], idx_o[, c("date", "vwret")], 
             by = "date")
rm(idx_o)
```

Merge the data sets

```{r merge}
library(data.table)
daily <- data.table(daily)
stopifnot(all(daily$date %in% idx$date))
mer <- merge(daily, idx, by = "date")
setkey(mer, permno, ym, date)

# clean up 
rm(daily)
```

# Estimate standard deviation from rolling regression
As in @Shumway01, we compute the idiosyncratic-risk by regressing
the stock return on the index and save the estimated standard deviation. We 
also save the $R^2$s and $\beta$s to check if the results make sense.

```{r comp_sigma}
# Get non-NA return columns
mer_sub <- mer[!is.na(ret), ]
setkey(mer_sub, permno, ym, date)
nrow(mer) - nrow(mer_sub)

func <- function(SD){
  min_obs <- 20L * 3L
  if(nrow(SD) < min_obs){
    fail_out <- rep(NA_real_, nrow(SD))
    fail_out <- list(fail_out, fail_out, fail_out)
    return(fail_out)
    
  }
  
  require(rollRegres)
  out <- suppressWarnings(roll_regres.fit(
    x = cbind(1, SD$vwret), y = SD$ret, width = 12L, 
    do_compute = c("sigmas", "r.squareds"),
    do_downdates = TRUE, grp = SD$ym, min_obs = min_obs))
  
  if(is.null(out))
    return(out)
  
  list(out$coefs[, 2], out$sigmas, out$r.squareds)
}
invisible(mer_sub[, c("beta", "sigma", "r_square") := func(.SD), by = permno])

# keep only the most recent figure each month
setkey(mer_sub, permno, ym, date)
invisible(mer_sub[, is_last := 1:.N == .N, by = .(permno, ym)])
mer_sub <- mer_sub[is_last == TRUE]
```

Check if the result is as expected. We start with plots

<div class="hideable">

```{r check_regress_res}
local({
  #####
  # number of observations
  nobs <- mer_sub[, .(nobs = sum(!is.na(sigma))), by = ym]
  nobs[, date := make_ym_inv(ym)]
  plot(nobs$nobs ~ nobs$date)
  
  #####
  # means
  tmp <- mer_sub[, .(
    beta     = mean(beta    , na.rm = TRUE), 
    sigma    = mean(sigma   , na.rm = TRUE), 
    r_square = mean(r_square, na.rm = TRUE)), by = ym]
  
  tmp[, date := make_ym_inv(ym)]
  par(mfcol = c(2, 2), mar = c(4.5, 4.5, .5, .5))
  plot(tmp$beta ~ tmp$date)
  plot(tmp$sigma ~ tmp$date)
  plot(tmp$r_square ~ tmp$date)
  
  #####
  # standard deviations
  tmp <- mer_sub[, .(
    beta     = sd(beta    , na.rm = TRUE), 
    sigma    = sd(sigma   , na.rm = TRUE), 
    r_square = sd(r_square, na.rm = TRUE)), by = ym]
  
  tmp[, date := make_ym_inv(ym)]
  par(mfcol = c(2, 2), mar = c(4.5, 4.5, .5, .5))
  plot(tmp$beta ~ tmp$date)
  plot(tmp$sigma ~ tmp$date)
  plot(tmp$r_square ~ tmp$date)
  
  #####
  # some high quantile
  tmp <- mer_sub[, .(
    beta     = quantile(beta    , na.rm = TRUE, probs = .9), 
    sigma    = quantile(sigma   , na.rm = TRUE, probs = .9), 
    r_square = quantile(r_square, na.rm = TRUE, probs = .9)), by = ym]
  
  tmp[, date := make_ym_inv(ym)]
  par(mfcol = c(2, 2))
  plot(tmp$beta ~ tmp$date)
  plot(tmp$sigma ~ tmp$date)
  plot(tmp$r_square ~ tmp$date)
})
```

</div>

Then overall summary statistics

```{r check_regress_res_stats}
local({
  for(x in c("beta", "sigma", "r_square")){
    z <- mer_sub[[x]]
    msg <- sQuote(x)
    cat(msg, "\n", rep("-", nchar(msg)), "\n", sep = "")
    print(summary(z[!is.na(z)]))
    cat("\n")
  }
})
```

Compare with summary statistics in @Shumway01 [pp. 114] (though he uses 
monthly data), @Campbell08 [pp. 2907], and @Duan12 [pp. 196]. 

We only need the standard deviations so we drop the other variables

```{r sigma_drop other}
invisible(mer_sub[, `:=`(
  date = NULL, mv = NULL, ret = NULL, market_size = NULL, vwret = NULL, 
  beta = NULL, r_square = NULL, is_last = NULL)])
sigma_vals <- mer_sub[!is.na(sigma)]
rm(mer_sub)
```

## Relative marked value

We compute the relative marked value in this section

```{r rel_marked_value}
mer_sub <- mer[, .(permno, ym, date, mv, market_size)]
setkey(mer_sub, permno, ym, date)

func <- function(mv, idx){
  # find the most recent non-na MV. Works if data is sorted
  latest_non_na <- which.max(which(!is.na(mv)))
  if(length(latest_non_na) == 0L)
    return(rep(NA_real_, length(mv)))
  rep(log(mv[latest_non_na]) - log(idx[latest_non_na]), length(mv))
}
invisible(mer_sub[, rel_size := func(mv, market_size), by = .(permno, ym)])

# only keep one observation per month and remove some variables
rel_size <- mer_sub[, is_last := .N == 1:.N, by = .(permno, ym)][
  is_last == TRUE][
    , `:=`(is_last = NULL, mv = NULL, date = NULL, market_size = NULL)]
rm(mer_sub)
```

Check result

<div class="hideable">

```{r rel_marked_value_check}
local({
  tmp <- copy(rel_size)
  tmp[, date := make_ym_inv(ym)]
  
  t1 <- tmp[, .(`frac na` = mean(is.na(rel_size))), by = date]
  with(t1, plot(`frac na` ~ date))
  
  tmp <- tmp[!is.na(rel_size)]
  t1 <- tmp[, .(mean = mean(rel_size), sd = sd(rel_size)), by = date]
  with(t1,{
    plot(mean ~ date)
    plot(sd   ~ date)
  })

  cat("summary\n")  
  print(summary(tmp$rel_size))
  
  cat("sd\n")
  print(sd(tmp$rel_size))
})
```

</div>

Compare with summary statistics in @Campbell08 [pp. 2907], 
@Chava11 [pp. 1272], and @Duan12 [pp. 196]. 

## Excess returns

We load in the monthly data and keep the variables we need

<div class="hideable">

```{r load_n_prep_monthly}
monthly <- readRDS(fs$monthly)
monthly <- subset(monthly, permno %in% permno_keep)

monthly <- data.table(monthly)
invisible(monthly[, `:=`(ret_log = log1p(ret), ym = make_ym(date))])
stopifnot(all(monthly[, length(unique(date)), by = .(permno, ym)]$V1 == 1L))
monthly <- monthly[, c("cusip", "permno", "gvkey", "ym", "ret_log")]

# check plot
local({
  tmp <- monthly[, .(mean = mean(ret_log, na.rm = TRUE)), by = ym]
  plot(tmp$mean ~ make_ym_inv(tmp$ym), type = "h")
})
```

</div>

We compute the log monthly returns for the index data

<div class="hideable">

```{r compute_idx_log_return}
# compute log return 
idx <- data.table(idx)
setkey(idx, date)
invisible(idx[
  , ym := make_ym(date)][
    , `:=`(is_last =  1:.N == .N, 
           vwret_log_m = sum(log1p(vwret))), 
    by = ym])
idx_month <- idx[is_last == TRUE][, is_last := NULL]
idx$is_last <- NULL
setkey(idx_month, ym)

setnames(idx_month, "vwret_log_m", "vwret_log")

# plot monthly return of index
par(mfcol = c(1, 2))
with(idx_month, plot(vwret_log ~ make_ym_inv(ym), type = "h"))
with(idx_month, 
     plot(cumsum(vwret_log[-1]) ~ make_ym_inv(ym[-1]), type = "l"))

# merge the two
monthly <- merge(monthly, idx_month[, .(ym, vwret_log)], by = "ym")
```

</div>

Compute average annualized difference

```{r comp_excess_ret}
func <- function(log_diff, ym, min_obs = 3L, width = 12L){
  out <- rep(NA_real_, length(log_diff))
  
  if(length(ym) < min_obs)
    return(out)
  
  ym <- ym - min(ym) + 1L # set `min(ym)` to one
  start_val <- min(which(ym >= min_obs))
  for(i in start_val:length(ym)){
    y <- ym[i]
    keep <- which(ym %in% (y - width + 1L):y)
    if(length(keep) < min_obs)
      next
    x <- log_diff[keep]
    x <- x[!is.na(x)]
    if(length(x) < min_obs)
      next
    out[i] <- mean(x) * 12 # annualize
  }
  
  out
}
setkey(monthly, permno, ym)
invisible(monthly[, excess_ret := func(ret_log - vwret_log, ym), by = permno])
```

Check result

<div class="hideable">

```{r comp_excess_ret_check}
local({
  #####
  # fraction of missing observations
  tmp <- monthly[, .(`frac na` = mean(is.na(excess_ret))), by = ym]
  par(mfcol = c(1, 2))
  with(tmp, {
    plot(`frac na` ~ make_ym_inv(ym))
    plot(`frac na` ~ make_ym_inv(ym), ylim = c(0, .15))
    })
  
  #####
  # mean, sd, and some high quantile
  tmp <- monthly[, .(mean =     mean(excess_ret, na.rm = TRUE), 
                     sd   =       sd(excess_ret, na.rm = TRUE),
                     q95  = quantile(excess_ret, na.rm = TRUE, probs = .95)), 
                 by = ym]
  par(mfcol = c(2, 2), mar = c(4.5, 4.5, .5, .5))
  with(tmp, {
    plot(mean ~ make_ym_inv(ym))
    plot(  sd ~ make_ym_inv(ym))
    plot( q95 ~ make_ym_inv(ym))
  })
  
  #####
  # summary stats
  cat("Summary stats before winsorizing\n")
  tmp <- monthly[!is.na(excess_ret), excess_ret]
  print(summary(tmp))
  cat("sd", sd(tmp), "\n\n")
  
  cat("Summary stats after winsorizing\n")
  tmp <- pmin(pmax(tmp, quantile(tmp, .05)), quantile(tmp, .95))
  print(summary(tmp))
  cat("sd", sd(tmp), "\n\n")
})
```

</div>

Compare with @Shumway01 [pp. 114] (though he does not have firms listed on NASDAQ), 
@Campbell08 [pp. 2907], and @Chava11 [pp. 1272]. 

## Merge data sets

We start by merging with a outer join 

```{r merge_dat}
rm(mer, idx, idx_month) # we do not need these anymore

monthly <- monthly[, c("ym", "cusip", "permno", "gvkey", "excess_ret")]
mer <- merge(monthly, sigma_vals, by = c("permno", "ym"), all = TRUE)
mer <- merge(mer    , rel_size  , by = c("permno", "ym"), all = TRUE)
c(nrow(monthly), nrow(sigma_vals), nrow(rel_size), nrow(mer))
```

We have some missing values and some that lag identifies. The latter is due 
to the outer join 

```{r check_frac_missing}
mer[, lapply(.SD, function(x) mean(is.na(x))), 
    .SD = c("excess_ret", "sigma", "rel_size", "gvkey")]

# we only want to keep those where we can match 
mer <- mer[!is.na(gvkey)]
mer[, lapply(.SD, function(x) mean(is.na(x))), 
    .SD = c("excess_ret", "sigma", "rel_size", "gvkey")]
nrow(mer)
length(unique(mer$permno))
```

We only have maps to `gvkey`s through the monthly data. We have not thrown away 
any of the `gvkey` year months as shown below

```{r no_throw_gvkey_ym}
local({
  # load monthly data and find those that will be included
  map_obj <- readRDS(fs$map_file)
  gvkey_keep <- unique(map_obj$gvkey)
  gvkey_keep <- subset(readRDS(fs$monthly), gvkey %in% gvkey_keep, 
                       select = c(gvkey, date))
  gvkey_keep$ym <- make_ym(gvkey_keep$date)
  
  # merge with mer, adding a dummy which we can check if any are NA after
  tmp <- merge(
    gvkey_keep, 
    with(mer, data.frame(gvkey, ym, is_mer = TRUE)), 
    by = c("gvkey", "ym"), all = TRUE)
  stopifnot(!anyNA(tmp$is_mer))
})
```

Save result

```{r save_res}
saveRDS(as.data.frame(mer), file.path("data", "market-datat.RDS"))
```

# References